\subsection[Binary Insertion Sort]{Binary Insertion Sort\cite{gfg_binaryinsertionsort}}
    \subsubsection{Overview}
        \begin{itemize}
            \item \textbf{History and Origin:} Binary insertion sort is a well-known variant of insertion sort that uses binary search to determine the correct insertion position. This technique was mentioned by John Mauchly in 1946 in one of the earliest published discussions on computer sorting\cite[p.~82]{knuth1998sorting}.
            \item \textbf{Efficiency:} Best for small datasets or nearly sorted/sorted data. Worst for reverse-sorted datasets.
            \item \textbf{Application:} Instead of using insertion sort when the subarray size becomes small in hybrid algorithms such as Timsort, Introsort, or Quicksort, binary insertion sort can be used for better efficiency. It is also useful in problems where the cost of comparisons is relatively high.
        \end{itemize}
    \subsubsection{Core Concepts}
        \begin{itemize}
            \item The array is divided into two parts: a sorted section and an unsorted section.
            \item For each element in the unsorted section, use \textbf{binary search\cite{gfg_binarysearch}} to find the correct position in the sorted section.
            \item Shift the elements to create space and insert the element into its correct position.
            \item After each insertion, the sorted section increases by one element.
        \end{itemize}

    \subsubsection{Step-by-step Expanation}
        \begin{itemize}
            \item Binary insertion sort uses the same code structure as Insertion Sort, except for the process of finding the correct position. Therefore, only this step is described below.
            \item Find the correct position \textit{pos} in $a[0\ldots (i - 1)]$ to insert a[i], where a[pos - 1] $\leq$ a[i] < a[pos].
            \item \textbf{Step 1:} Set val = a[i], \textit{L} = 0, \textit{R} = i - 1 and \textit{pos} = i.
            \item \textbf{Step 2:} While L $\leq$ R.
                \begin{itemize}
                    \item Set Mid = (L + R) / 2.
                    \item If a[Mid] $\geq$ val then set R = Mid - 1 and pos = Mid.
                    \item Else set L = Mid + 1.
                \end{itemize}
            \item \textbf{Step 3:} Move forward a[pos \ldots (i - 1)] to one element. After that, set a[pos] = val. 
        \end{itemize}

    \subsubsection{Pseudo Code}
        \begin{itemize}
            \item \textbf{Algorithm:} Binary Insertion Sort
            \item \textbf{Input:} An array A of n elements
        \end{itemize}
        \begin{verbatim}
            1. for i = 1 to length(A)-1
            2.      value = A[i]
            3.      l = 0, r = i-1, pos = i
            4.      while l <= r
            5.           m = l + (r-l) / 2
            6.           if A[m] >= value
            7.                pos = m
            8.                r = m-1
            9.           else
            10.               l = m+1
            11.     for j = i downto pos+1
            12.          A[j] = A[j-1]
            13.     A[pos] = value
        \end{verbatim}
    \subsubsection{Complexity Analysis}
        \begin{itemize}
            \item \textbf{Time Complexity:}
            \begin{itemize}
                \item Finding correct position takes $O(logn)$ but the process of insertion still takes $O(n)$.
                \item Best case: $O(nlogn)$
                    \subitem When the input is already sorted.
                \item Average case: $O(n^2)$
                    \subitem When the input is random.
                \item Worst case: $O(n^2)$ 
                    \subitem When the input is reverse-sorted.
            \end{itemize}
            \item \textbf{Space Complexity:} $O(1)$ 
                \subitem In-place algorithm. Requires only a constant amount of extra memory.
        \end{itemize}
        
    \subsubsection{Variants and Optimizations}
        \begin{itemize}
            \item \textbf{Clustered Binary Insertion Sort:\cite{goel2018cbis}}
            \begin{itemize}
                \item A variant of Binary Insertion Sort.
                \item Uses the concept of clusters to determine the range for searching the correct insertion position.
                \item By using a previously updated position pointer (POP), the algorithm can determine which cluster to perform the binary search on, making it slightly faster than BIS in some cases.
            \end{itemize}
        \end{itemize}