\subsection[Heap Sort]{Heap Sort \cite{williams1964heapsort}}

\subsubsection{Overview}
\begin{itemize}
    \item \textbf{Origin:} Heapsort was invented by J. W. J. Williams in 1964.
    \item \textbf{Efficiency:} Heap sort is a highly efficient comparison-based sorting algorithm that utilizes a binary heap structure.
    \item \textbf{Application:} Heap sort can be applied to large datasets due to its consistent running time of $O(nlogn)$ for all cases.
\end{itemize}

\subsubsection{Core Concepts}
\begin{itemize}
    \item Heap sort is an improved version of Selection sort, which operates based on the principle of building a Max-Heap from unsorted elements to sort arrays in ascending order.
    \item The root of the heap is always the largest element and it will be placed in its first correct position in the sorted portion.
\end{itemize}

\subsubsection{Step-by-step Explanation}
\begin{itemize}
    \item Assume an array of $n$ elements: $a[0 \dots (n-1)]$.
    \item The algorithm consists of two main stages:
    
    \item \textbf{Stage 1: Heap construction}
    \begin{itemize}
        \item Construct a binary heap from elements indexed from $0$ to $n-1$.
        \item Rearrange elements so that the largest element becomes the root of the heap $($ the element at index 0$ )$.
    \end{itemize}

    \item \textbf{Stage 2: Sorting}
    \begin{itemize}
        \item Swap the root element with the last element in the unsorted portion.
        \item Reduce the heap size by 1 and repeat the process at step 2 with the new heap size
    \end{itemize}

    \item The algorithm terminates when all elements are sorted.
\end{itemize}

\subsubsection{Pseudo code}

\begin{verbatim}
    1. Heapify(A, start, n)
    2.  leftChild = start * 2 + 1
    3.  rightChild = start * 2 + 2
    4.  LargerChild = leftChild
    5.  if leftChild >= n: return
    6.  if A[leftChild] < A[rightChild]:
    7.      LargerChild = rightChild
    8.  if A[start] < A[LargerChild]:
    9.      swap(A[start], A[LargerChild])
    10.      Heapify(A,LargerChild,n) 
\end{verbatim}

\begin{verbatim}
1. BuildMaxHeap(A,n)
2. for i = n-1 down to 1:
3.     swap(A[0], A[i])
4.     heapSize = heapSize - 1
5.     Heapify(A, 0, heapSize)
\end{verbatim}



\subsubsection{Complexity Analysis}
\begin{itemize}
    \item \textbf{Best case:} $O(n \log n)$
    \item \textbf{Average case:} $O(n \log n)$
    \item \textbf{Worst case:} $O(n \log n)$
\end{itemize}

\subsubsection{Variants and Optimizations}
\begin{itemize}
    \item Variants of this algorithm is the adjustment for the number children.
    \item \textbf{Current implementation:  }Binary heap: Build Max-Heap with each node has at most two children $( $ Left and Right$)$
    \item \textbf{Another implementation: }Ternary heap \cite{johnson1975priority}: each node has at most three children, which reduce swaps but increase comparisons
\end{itemize}