\subsection[Radix Sort]{Radix Sort \cite{shun2019parallel}}
    \subsubsection{Overview}
        \begin{itemize}
            \item \textbf{History and Origin:} A non-comparative sorting algorithm originally used for sorting punch cards, with early implementations on electronic computer systems dating back to 1956 \cite{friend1956sorting}.
            \item \textbf{Efficiency:} Highly efficient for datasets with keys that are not too large.
            \item \textbf{Application:} Widely used where numerical data needs to be sorted in linear time \cite{cormen2022introduction}.
        \end{itemize}
    \subsubsection{Core Concepts}
        \begin{itemize}
            \item Radix Sort avoids direct comparison between elements, instead grouping them by their individual digits.
            \item The implemented version is Least Significant Digit (LSD) Radix Sort (Base-10).
            \item It uses 10 structural \texttt{Buckets}, each containing a dynamically allocated array to directly store and retrieve elements for each digit from 0 to 9.
        \end{itemize}

    \subsubsection{Step-by-step Explanation}
        \begin{itemize}
            \item \textbf{Step 1:} Find the maximum value (\textit{maxv}) in the array to determine the number of digits.
            \item \textbf{Step 2:} Initialize 10 \texttt{Buckets}, dynamically allocating an array of size $n$ for each bucket's data.
            \item \textbf{Step 3:} Initialize a multiplier \textit{j} = 1.
            \item \textbf{Step 4:} While \textit{maxv} / \textit{j} > 0:
            \begin{itemize}
                \item Reset the \textit{count} of all 10 buckets to 0.
                \item For each element, extract its digit: \textit{digit} = (array[i] / j) \% 10.
                \item Append the element to the corresponding bucket's \textit{data} array and increment its \textit{count}.
                \item Traverse the 10 buckets in order (0 to 9) and sequentially write the elements back into the original array.
            \end{itemize}
            \item \textbf{Step 5:} Multiply \textit{j} by 10 and return to \textbf{Step 4} until all digits are processed.
            \item \textbf{Step 6:} Deallocate the memory of the bucket arrays.
        \end{itemize}

    \subsubsection{Pseudo Code}
        \begin{itemize}
            \item \textbf{Algorithm:} LSD Radix Sort (Bucket Approach)
            \item \textbf{Input:} An array A of n elements
        \end{itemize}
        \begin{verbatim}
        1. Find maxv in A
        2. Initialize 10 Buckets (each has an array 'data' and 'count')
        3. j = 1
        4. While (maxv / j) > 0:
        5.    For k = 0 to 9:
        6.       Buckets[k].count = 0
        7.    For i = 0 to n - 1:
        8.       digit = (A[i] / j) % 10
        9.       empty = Buckets[digit].count
        10.      Buckets[digit].data[empty] = A[i]
        11.      Buckets[digit].count = Buckets[digit].count + 1
        12.   dex = 0
        13.   For k = 0 to 9:
        14.      For m = 0 to Buckets[k].count - 1:
        15.         A[dex] = Buckets[k].data[m]
        16.         dex = dex + 1
        17.   j = j * 10
        18. Free Bucket memory
        \end{verbatim}
    \subsubsection{Complexity Analysis}
        \begin{itemize}
            \item \textbf{Time Complexity:} $O(d(n + k))$
                \subitem Where \textit{d} is the number of digits, \textit{n} is the number of elements, and \textit{k} is 10.
            \item \textbf{Space Complexity:} $O(n \times k)$ 
                \subitem The current implementation allocates 10 full-sized arrays of size \textit{n}, requiring heavily expanded auxiliary memory.
        \end{itemize}

    \subsubsection{Variants and Optimizations}
        \begin{itemize}
            \item \textbf{Current Implementation (Base-10 with Dynamic Buckets):}
            \begin{itemize}
                \item The provided C++ code creates 10 explicit buckets, each allocating $O(n)$ memory space. This is intuitive but highly memory-intensive for large datasets.
            \end{itemize}
            \item \textbf{Theoretical Upgrade 1: Counting Sort Array (Standard LSD):}
            \begin{itemize}
                \item \textbf{Optimization:} Instead of 10 arrays of size $n$, use a single prefix-sum \texttt{Count} array of size 10 and one temporary output array of size $n$.
                \item \textbf{Benefit:} Reduces the Space Complexity drastically from $O(10n)$ to $O(n + 10)$.
            \end{itemize}
            \item \textbf{Theoretical Upgrade 2: Bitwise Radix Sort (Base-256):}
            \begin{itemize}
                \item \textbf{Optimization:} Change the radix base to a power of 2 (e.g., base 256) using fast bitwise operations (\texttt{>>} and \texttt{\&}).
                \item \textbf{Benefit:} Eliminates slow division (\texttt{/}) and modulo (\texttt{\%}) operations, executing much faster at the CPU level.
            \end{itemize}
            \item \textbf{Theoretical Upgrade 3: In-Place Radix Sort \cite{shun2019parallel}:}
            \begin{itemize}
                \item \textbf{Optimization:} Modify the permutation phase to run in-place by tracing cycles.
                \item \textbf{Benefit:} Eliminates auxiliary array overhead entirely, massively improving cache hits.
            \end{itemize}
        \end{itemize}