\subsection[Shaker Sort]{Shaker Sort\cite{gfg_cocktailsort}}
    \subsubsection{Overview}
        \begin{itemize}
            \item \textbf{History and Origin:} A variant of the bubble sort algorithm, also known as cocktail sort, cocktail-shaker sort, bidirectional bubble sort, etc.
            \item \textbf{Efficiency:} Best for small datasets or nearly sorted/sorted data. Worst for reverse-sorted datasets.
            \item \textbf{Application:} Used for better efficiency than bubble sort in some cases. It is also used for teaching basic sorting algorithms or for sorting small datasets.
        \end{itemize}
    \subsubsection{Core Concepts}
        \begin{itemize}
            \item Shaker sort compares adjacent elements and swaps their positions if necessary.
            \item The algorithm moves larger elements to the end of the array and smaller elements to the beginning of the array. These movements are similar to bubbles floating to the surface and sinking to the bottom.
        \end{itemize}
    \subsubsection{Step-by-step Explanation}
        \begin{itemize}
            \item Assume an array of $n$ elements, $a[0 \ldots (n-1)]$.
            \item \textbf{Step 1:} Set \textit{swapped} = true.
            \item \textbf{Step 2:} While \textit{swapped} is true, repeat the following steps.
            \item \textbf{Step 3:} Set \textit{swapped} = false. Run a loop from $0$ to $n-2$. If an adjacent pair is not in ascending order, swap them and set \textit{swapped} = true.
            \item \textbf{Step 4:} If \textit{swapped} = false, stop the algorithm.
            \item \textbf{Step 5:} Set \textit{swapped} = false. Run a loop from $n-2$ down to $0$. If an adjacent pair is not in ascending order, swap them and set \textit{swapped} = true.
            \item \textbf{Step 6:} Repeat \textbf{Step 2}.
        \end{itemize}
    \subsubsection{Pseudo Code}
        \begin{itemize}
            \item \textbf{Algorithm:} Shaker Sort
            \item \textbf{Input:} An array A of n elements
        \end{itemize}
        \begin{verbatim}
            1.  swapped = 1
            2.  while swapped == 1
            3.       swapped = 0
            4.       for i = 0 to n-2
            5.            if A[i] > A[i+1]
            6.                 swap A[i], A[i+1]
            7.                 swapped = 1
            8.
            9.       if swapped == 0
            10.           break
            11.
            12.      swapped = 0
            13.      for i = n-2 downto 0
            14.           if A[i] > A[i+1]
            15.                swap A[i], A[i+1]
            16.                swapped = 1
        \end{verbatim}
    \subsubsection{Complexity Analysis}
        \begin{itemize}
            \item \textbf{Time Complexity:}
            \begin{itemize}
                \item \textbf{Best case:} $O(n)$
                    \subitem When the input is already sorted.
                \item \textbf{Average case:} $O(n^2)$
                    \subitem When the input is random.
                \item \textbf{Worst case:} $O(n^2)$ 
                    \subitem When the input is reverse-sorted.
            \end{itemize}
            \item \textbf{Space Complexity:} O(1) 
                \subitem In-place algorithm. Requires only a constant amount of extra memory.
        \end{itemize}
        
    \subsubsection{Variants and Optimizations}
        \begin{itemize}
            \item \textbf{Past implementation:\cite{iverson1962apl}}
                \begin{itemize}
                    \item An early implementation related to Shaker Sort.
                    \item Compares all adjacent pairs of elements once in each iteration but traverses the array with forward and backward passes sequentially.
                    \item If a pair is swapped, the algorithm updates the array to save information about whether those elements are already in their correct relative position, so they can be skipped in future comparisons.
                \end{itemize}

            \item \textbf{Original implementation: } Original shaker sort
            \begin{itemize}
                \item Compares all adjacent pairs of elements twice in each iteration: one forward pass and one backward pass.
                \item Optional optimization: After each iteration, decrease the range by increasing \textit{start} and decreasing \textit{end}, since the elements at the beginning and the end are already in their correct positions.
                \item Use an additional flag; if no swaps occur in an iteration, the algorithm stops early.
            \end{itemize}

            \item \textbf{Improved implementation:\cite[Slide~31]{thao_sorting_slides}} Shaker sort with last swapped position
            \begin{itemize}
                \item Keeps the same structure as the original shaker sort but adds the following optimization.
                \item An additional variable \textit{pos} stores the position of the last swap in the array.
                \item We only need to consider the range up to \textit{pos}, because elements outside this range are already in their correct positions. This helps reduce the effective range of the array.
            \end{itemize}
        \end{itemize}