\subsection[Shell Sort]{Shell Sort \cite{afzal2007enhancedshellsort}}
    \subsubsection{Overview}
        \begin{itemize}
            \item \textbf{Origin: } The shell sort is a popular variant of insertion sort. It is first published in 1959 by Donald Shell.
            \item \textbf{Efficiency: } This algorithm is appropriate for medium-sized and sorted/nearly sorted datasets.
            \item \textbf{Application: } The shell sort can be applied in reality to deal with medium-sized data to replace other sorting algorithms namely Merge sort, which are too heavy. Sometimes, this technique can be used as a subroutine in hybrid algorithms. 
        \end{itemize}
    \subsubsection{Core concepts}
        \begin{itemize}
            \item The shell sort uses the same operating principles with insertion sort. However, the key difference is that Shell sort instead of comparing adjacent elements to find place for the insertion, it compares elements which are far apart.
        \end{itemize}
    \subsubsection{Step-by-step Expanation}
    \begin{itemize}
        \item Assume that the array of n elements, a[0...(n-1)].
        \item \textbf{Step 1: } Set the gap base on a certain sequence , usually $(n/2)$
        \item \textbf{Step 2: } Divide the array into several sublist and do insertion sort in those.
        \item \textbf{Step 3: } Reduce the gap and repeat the process on \textbf{Step 2} if the gap is greater than 1.
        \item \textbf{Step 4:} If the gap equals to 1, do the last insertion sort when the array is nearly sorted.
    \end{itemize}
    \subsubsection{Pseudo code} 
        \begin{verbatim}
            1.  gap = n/2
            2.  while gap >= 1:
            3.      for i = gap to n-1:
            4.          key = a[i]
            5.          j = i - gap
            6.          while (j >= 0 and a[j] > key):
            7               a[j + gap] = a[j]
            8.              j = j - gap
            9.          a[j + gap] = a[j]
            10.         gap = gap / 2      
        \end{verbatim}

    \subsubsection{Complexity Analysis}
        \begin{itemize}
            \item Best case: $O(nlogn)$
            \subitem If choose the suitable sequence for the array.
            \item Worst case: $O(1.5n)$
            \subitem If choose the wrong sequence for the problem.
            \item Average case: $O(n^{1.25}) $
        \end{itemize}
    

    \subsubsection{Variants and Optimizations} 
    \begin{itemize}
        \item Variants of Shell sort is mainly the choose of sequence:
        \begin{itemize}
            \item Knuth's Increments: $1,3,13,...,(3k-1)/2$
            \item Sedgewick's Increments: $1,8,23,77,...$
            \item Hibbard's Increments: $1,3,7,15,31,...$
        \end{itemize}
        \item Each Optimization has its own pros and cons. Depend on the input data, we can choose a distinct version to achieve best performance
    \end{itemize}
